Taking Linear Logic Apart

### 2019-06-03 doc/paper/main.tex
Rewrote section explaining why no hyperenvironments for logical rules

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piRecv{x}{}{\piBoundSend{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \\[1\baselineskip]
  {Structural rules}
  \begin{center}
    \hcpInfAx
    \hcpInfCut
  \end{center}
  \begin{center}
    \hcpInfMix
    \hcpInfHalt
  \end{center}
  {Logical rules}
  \begin{center}
    \hcpInfBoundTens
    \hcpInfParr
  \end{center}
  \begin{center}
    \hcpInfOne
    \hcpInfBot
  \end{center}
  \begin{center}
    \hcpInfPlus1
    \hcpInfPlus2
  \end{center}
  \begin{center}
    \hcpInfWith
  \end{center}
  \begin{center}
    \hcpInfNil
    \hcpInfTop
  \end{center}
\end{definition}\noindent

Note that the logical rules disallow hyperenvironments.

>>>>

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piRecv{x}{}{\piBoundSend{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \\[1\baselineskip]
  {Structural rules}
  \begin{center}
    \hcpInfAx
    \hcpInfCut
  \end{center}
  \begin{center}
    \hcpInfMix
    \hcpInfHalt
  \end{center}
  {Logical rules}
  \begin{center}
    \hcpInfBoundTens
    \hcpInfParr
  \end{center}
  \begin{center}
    \hcpInfOne
    \hcpInfBot
  \end{center}
  \begin{center}
    \hcpInfPlus1
    \hcpInfPlus2
  \end{center}
  \begin{center}
    \hcpInfWith
  \end{center}
  \begin{center}
    \hcpInfNil
    \hcpInfTop
  \end{center}
\end{definition}
Note that the logical rules disallow hyperenvironments. If they did not, the following processes would be well-typed, even though they are stuck:
\begin{center}
  \begin{prooftree*}
    \noLine\AXC{$\vphantom{\Gamma}$}\UIC{$\vphantom{\Gamma}$}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [\piHalt]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{z}{}{\piHalt}]
      {\tmty{z}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{x}{(\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}})}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}\hspace*{-1.75em}%
  \begin{prooftree*}
    \AXC{}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [{\piHalt}]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{y}{}{\piHalt}]
      {\tmty{y}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{y}{\one}}$}
    \AXC{}
    \NOM{Ax}
    \UIC{$\seq
      [\cpLink{x}{z}]
      {\tmty{x}{\bot}, \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{y}{}{\cpLink{x}{z}}]
      {\tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Mix}
    \BIC{$\seq
      [(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})]
      {\tmty{x}{\one} \hsep \tmty{y}{\one} \hsep \tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{x}{\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}
\end{center}
The first process is self-locking, in that it holds both endpoints of a channel in the same process, and hence communication can never happen. Self-locking processes can be ruled out by requiring that the two endpoints of a channel are held in different processes.

The second term is more complicated: it is not a self-lock, and not a traditional deadlock, where two processes are waiting on each other. However, it is definitely stuck. The problem lies with readiness: a process typed under a hyperenvironment $\ty{\Gamma_1}\hsep\dots\hsep\ty{\Gamma_n}$, it promises to act as, essentially, $n$ independent processes, each of which ready to act on a channel from $\ty{\Gamma_i}$. However, the process $\tm{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}$, a single process, can be typed under the hyperenvironment $\ty{\tmty{x}{\one} \hsep \tmty{y}{\one}}$, which consists of \emph{two} environments. We can abuse this failure of readiness by asking the process to communicate on $\tm{y}$, which it is incapable of doing immediately. This is not a property of the $(\one)$ rule allowing hyperenvironments. For instance, a similar example can be constructed using the $(\tens)$ rule.
